Kanwal (2017) BDA Notebook

### 2018-06-13 analysis/BDA-analysis.Rmd
re-organize dirs

```{r packages, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(rwebppl)
library(tidyr)
```

@@

# BDA - Modeling participant data

>>>>

[Kanwal (2017)](https://www.ncbi.nlm.nih.gov/pubmed/28494263) present a test of Zipf's `Law of Abbreviation` -- that a universal design feature of human language emerges as a result of individuals optimising form-meaning mappings under competing pressures to communicate accurately (`expressivity`), but also efficiently (roughly, `learnabilty`). Their study uses miniature artificial language learning paradigm. They show that language users optimise form-mean mapping only when pressures for accuracy and effficency *both* operate during the communicative task.

In this analysis we explore the degree to which the observed behavior can be explained by more general purpose social reasoning mechanisms using a computational pragmatics model. We show that our model accounts for a large portion of the variance ($r^2=0.97$) across multiple conditions (both optimal and sub-optimal behavior).

While one interpretation of this finding frames the results as entirely pragmatics-driven. Instead we argue that the Zipfian result is a bi-product of pragmatic mechanims.

Summary of paper:

Four conditions:

  * `combination`: both communicative (expressivity) and production cost (efficiency) pressures.

  * `accuracy`: only communicative pressure (no production cost).

  * `timing`: no communicative pressures (only production costs).

  * `neither`: not communicative pressures or productino cost. Note this scenario is not very interesting to us as it does not resemble any aspects of communication.

  ![Kanwal (2017), fig3](img/kanwal-2017-fig3.png)

```{r packages, message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(rwebppl)
library(tidyr)
```

@@

# BDA - Modeling Zipf's Law of Abbreviation via Kanwal (2017)

Here we model participant behavior as pragmatic reasoning instantiated in the Rational speech-act family of models. We infer four hyperparmeters which encode different participant strategies (optimal in different conditions).

### BDA generative model

* `semantics`: $Unif(["natural", "none"])$ where
    * $natural: [[zop]] \in \{zopekil, zopudon\}, [[zopudon]] \in \{zopudon\}, [[zop]] \in \{zopekil\} \in \{zopekil\}$
    * $none: [[x]] \in \{zopekil, zopudon\}$

* `costs`: $Unif(["present", "not-present", "reversed"])$ where
    * $present: "zop":0.3, "zopudon":0.7, "zopekil":0.7$
    * $present: cost(zop) = cost(zopekil) = cost(zopudon) = 0$
    * $reversed: "zop":0.7, "zopudon":0.3, "zopekil":0.3$

* `recursion-level`: $Categorical([1, 2])$

* `needProbabilities`: $Categorical([present, not present, reversed])$
    * `present`: "zopudon":24,"zopekil":8,  
    * `not-present`: "zopudon":1,"zopekil":1,
    * `reversed`: "zopudon":8,"zopekil":24,

### Strategy mapping to hyperparams

Three conditions of interest:

  * `combination`:  `semantics`='natural', `costs`='present', `needProbabilities`='present'

  * `accuracy`:  `semantics`='natural', `costs`='not-present'

  * `timing`:  `semantics`='not-present', `costs`='present'

## Analysis
